### 2024-01-25 article.Rmd
anonymisation

---
title: "(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games"
# author:
#   "Ismail Guennouni^1^, Samuel Dupret^1^, Quentin JM Huys^2^, Maarten Speekenbrink^1^"

keywords:
- Social Trust
- Repair of Cooperation
- Cognitive Intervention 
- Economic Games 
- Hidden Markov Models

abstract: |
  Social trust is a key component of strong social bonds. Its absence is a risk factor for social dysfunction and psychopathology. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. We used the Repeated Trust Game to evaluate a cognitive intervention aiming to repair cooperation breakdowns caused by a one-time defection by the co-player. In this study, 318 participants played the role of trustee across two online games, believing they were facing two different players. Between games, some received a brief cognitive intervention. Post-intervention, participants showed more cooperative behavior both before and after defection by the co-player. Analysis using a hidden Markov model revealed that those in the intervention group had a higher proportion of cooperative states than participants in the control group. This is consistent with participants inferring from the intervention that pro-social behavior may generally provide more beneficial outcomes to them in the long run.

bibliography: "bib/cogIntervention.bib"
biblio-style: spphys

output:
  bookdown::pdf_document2:
    toc: false
header-includes:
  - \usepackage{lscape}
---
<!-- \small   -->
<!-- ^1^ *Department of Experimental Psychology, Division of Psychology and Language Sciences, UCL*.   -->
<!-- ^2^ *Division of Psychiatry and Max Planck Centre for Computational Psychiatry and Ageing Research, Queen Square Institute of Neurology, UCL.* -->

---

```{r setupCoax, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
knitr::opts_chunk$set(out.width = "\\textwidth")
```

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
#require(citr)
require(bookdown)

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
library(magick)

```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
options(tinytex.verbose = TRUE)
```

## Disclosures {.unnumbered}

Conflicts of interest: The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
Funding: Author 1 was supported by the UK Engineering and Physical Sciences Research Council under grant EP/S515255/1. Author 3 acknowledges support by the UCLH NIHR BRC. He has received fees and options for consultancies for Aya Technologies and Alto Neuroscience.
Ethics: This research received approval from a local ethics board (ID:21029/001).
Artificial intelligence: No artificial intelligence assisted technologies were used in this research or the creation of this article.
Reproducibility: The results are fully reproducible, and all data and code for the analyses (including the manuscript in R Markdown format) are available. The design and analysis plans were not preregistered.

<!-- Author Contribution: I. Guennouni, QJM. Huys and M. Speekenbrink designed and developed the study concept. Experiment design was done by S. Dupret and I. Guennouni. Testing and data collection were performed by I. Guennouni. I. Guennouni analysed and interpreted the data under the supervision of QJM. Huys and M. Speekenbrink. M. Speekenbrink and I. Guennouni jointly performed the HMM modelling. All authors jointly wrote and approved the final version of the manuscript for submission. -->

# Introduction

Cooperation refers to the collaborative efforts of individuals or entities working towards a shared goal or advantage [@tomasello2012two]. This coordinated approach often leads to outcomes that exceed the sum of individual contributions.  Trust is a key driver of cooperation, enabling individuals to engage in mutually beneficial interactions without fear of exploitation [@balliet2013trust]. This aligns with some definitions of trust as “a psychological state comprising the intention to accept vulnerability based on positive expectations of the intentions or behavior of another” [@rousseau1998not]. When we trust others, we are more likely to cooperate with them, even when there is a risk of being cheated. Prior research also emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016; @meng_multilevel_2014]. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders [@fonagy_role_2014].

A well-established paradigm in the study of trust is the Repeated Trust Game (RTG; @joyce_trust_1995). In this game, the "investor" decides how much of an endowment to send to the other player (the "trustee"). The amount that is sent is tripled and the trustee decides, in return, how much of the tripled amount to send back to the investor. The Nash equilibrium for a single-round version of the game is for the investor to send nothing. In the repeated version, rewards for both players are maximised if they build trust and share the benefits of higher investments. If the investor is rewarded for taking the risk of sending an investment, they are likely invest again in future rounds. But if the investor obtains a low return on their investment, they will likely reduce future investments, thereby diminishing the potential gains for both players.

Since a lack of epistemic trust is a risk factor for social dysfunction [@fonagy_mentalizing_2017], and given the importance of trust for building and maintaining strong social bonds, interventions that foster and strengthen trust-based cooperation would be highly beneficial to society. Such interventions would allow people to more easily repair broken relationships and continue harvesting the benefits of cooperation even in the presence of accidental or intentional social norm violations.

To encourage the emergence and maintenance of trust in the RTG setting, some studies have focused on modification of the game, e.g. through introducing a third-party who monitors the actions of the other players [@charness_investment_2008; @fiedler_effect_2017]. Others chose to prime participants towards feeling gratitude [@drazkowski_gratitude_2017] or with the concepts of friend and foe [@burnham_friend-or-foe_2000]. Whilst these results show that it is possible to improve cooperative outcomes at the start of the game, they do not address how to repair a breakdown of trust after intentional or accidental non-cooperative actions by the players. Cooperation in the RTG can easily break down after a transgressive behavior such as a nil or very low investment by the investor or a return by the trustee below the investment sent [@bendor_when_1991]. Such breakdowns of cooperation are particularly evident when the trustee suffers from social disorders such as Borderline Personality Disorder (BPD; @lieb_borderline_2004). Trustees with BPD fail to engage in trust-repairing behaviors such as coaxing the investor by signalling trustworthiness via sending high returns [@king-casas_rupture_2008].  Neuroimaging analysis from the study of @king-casas_rupture_2008 is consistent with the idea that individuals with BPD may not recognize how offering low returns in such interactions conflicts with societal expectations. Further, computational analysis by @hula2018model links this issue to some BPD individuals' mentalizing difficulties, such as their inability to foresee the impact of their low returns on the investor's future actions.

In devising interventions to repair trust in the RTG, we can take inspiration from cognitive interventions championed by successful therapies that aim to improve social dysfunction in BPD. Psychotherapies such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993] have been shown to improve social skills in BPD [@gunderson_borderline_2018]. However, response to these treatments is highly variable, and determining which interventions are effective for particular sub-groups of patients is challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific components of psychotherapeutic treatment affect quantitative markers of behavior such as those inferred through computational models [@huys_computational_2016; @reiter_neuro-cognitive_2021; @dercon2022core]. Combining the use of specific cognitive probes inspired by therapeutic interventions and computational models of behavior may allow us to uncover the cognitive mechanisms targeted by common forms of psychotherapy. In turn, this may provide the basis for choosing effective psychotherapeutic interventions for given individuals.

In this study, using a randomised control trial design, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation in the RTG from a low investment sent by a computerized investor. The intervention focuses on explaining the potential harm from reciprocating non-cooperative actions and suggesting a non-impulsive course of action to coax the investor back into cooperation. 
The computerized investor was programmed to play according to a hidden Markov model estimated from real players' data in prior research. A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control. To foreshadow our results, we find that the intervention led to more cooperative actions (higher returns) by the participants and countered a tendency to send back lower returns after a transgression from the investor. However, we found no evidence that the effects of the intervention transfer to a different (Repeated Prisoner’s Dilemma) game.

```{r dataLoad, include=FALSE }

d_finished <- readRDS("data/d_anonym.RDS")

sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE)
```

# Method

## Participants

A total of 318 participants were recruited on the Prolific Academic platform (prolific.co). The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years, with a `r round(sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years standard deviation. Participants were paid a fixed fee of £5 plus a bonus payment dependent on their performance that averaged £0.71.

## Design and Procedure

The experiment had a 2 (Condition: Intervention or Control) by 3 (Game: Trust-Game Pre Intervention, Trust-Game Post Intervention, Prisoner's Dilemma Post Intervention) design, with repeated measures on the second factor. Participants were randomly assigned to one of the two levels of the first factor. The games were designed and implemented online using Empirica [@almaatouq2021empirica].

```{r timeline, include=F, fig.cap = "Experiment overview. After playing 15 rounds of the RTG as the trustee, participants were randomized to either be part of the control or intervention condition. They then played the second set of 15 rounds of RTG (again as the trustee) to examine intervention effects, and 7 rounds of a repeated Prisoner’s Dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="100%", out.height= "120%"}

knitr::include_graphics("figures/timeline.png")

```

## Tasks and Measures

### Repeated Trust Game

Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \ref{fig:HMMPanels}.A). The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players.

The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \ref{fig:HMMPanels}.C). These distributions reflect "low-trust", "medium-trust", or "high-trust". Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.D). In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-intervention) or round 13 (post-intervention). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The initial state for the HMM investor in each instance of the game was the "mid-trust" state.

On each round, immediately after being informed of the investment sent, participants in the intervention condition were asked to provide an evaluation of their emotion in terms of valence (from negative to positive) and arousal (from low to high). Participants in the control condition were asked to evaluate the investment in terms of speed (from slow to fast) and magnitude (from low to high). These evaluations were made by clicking on a two-dimensional field with labelled axes as shown in Figure \ref{fig:HMMPanels}.C for the intervention condition.

### Repeated Prisoner's Dilemma

To ascertain whether any effect of the intervention would transfer to a different game, participants played 7 rounds of a Repeated Prisoner’s Dilemma (RPD). In each round, participants could choose between a cooperative action with a reward of 5 (the other player also cooperates) or 1 (the other player defects), or a defect action with a reward of 7 (the other player cooperates) or 2 (the other player defects). The Nash equilibrium for a single-round version is to choose the non-cooperative action. In the repeated version, both players can maximise their reward by choosing to cooperate.

```{r, include=FALSE}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

Neutral <- c(0.003393529, 0.006930874, 0.013053553, 0.022671228,
0.036310144, 0.053627606, 0.073039389, 0.091734929, 0.106248217,
0.113479701, 0.111769796, 0.101517390, 0.085028780, 0.065675083,
0.046778251, 0.030725245, 0.018610323, 0.010394861, 0.005354126,
0.002543094, 0.001113881)

Pro_social <- c(0.003162697, 0.001057162, 0.001410197, 0.001881016,
0.002508877, 0.003346113, 0.004462480, 0.005950950, 0.007935434,
0.010581067, 0.014107909, 0.018809194, 0.025075644, 0.033427845,
0.044559370, 0.059394206, 0.079163228, 0.105506029, 0.140606514,
0.187373417, 0.249680651)

investment <- seq(0:20) -1

response_probs <- as.data.frame(cbind(investment,Anti_social,Neutral,Pro_social)) %>% 
  rename("low-trust" = "Anti_social", "medium-trust"="Neutral", "high-trust" = "Pro_social") %>%
  pivot_longer(cols=c("low-trust","medium-trust","high-trust"),
                    names_to='Investor_state',
                    values_to='probability') %>% 
   mutate(across(Investor_state, factor, levels=c("low-trust","medium-trust","high-trust")))
```

```{r, include=FALSE}

plotinvHMM <- ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state', x = "Investment", y= "Probability") + 
  theme_bw() + 
  theme(legend.position = "bottom",  legend.text = element_text(size = 12),  legend.title = element_text(size = 14))

plotinvHMM
```

```{r, include=FALSE}

unhappy_pars <- rbind(c(3.5435248 , 0.0592584), c(0.1611541, 0.4675468), c(0,0)) 
neutral_pars <- rbind(c(0.8846036, - 0.4394031), c(2.4214322, - 0.0710711), c(0,0))
happy_pars <- rbind(c(-2.004988 ,- 0.151186), c(-1.2976976 , - 0.1357088), c(0,0))

pars_inv <- list(unhappy_pars, neutral_pars, happy_pars)

plot_HMM_transitions <- function(ns, pars_mat) {

  trans_prob <- data.frame(
    from = rep(1:ns, each=100*ns),
    to = rep(1:ns, each=100),
    ret = seq(-20,60,length=100),
    probs = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=100)

  for(from in 1:ns) {
  pars <- matrix(pars_mat[[from]], ncol=2)
  # print(pars)

    for(to in 1:ns) {
        x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"ret"]
        y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)

    for(to in 1:ns) {
      trans_prob$probs[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  df <- as.data.frame(trans_prob) %>% 
    mutate(from = recode(from, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust"),
           to = recode(to, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust") ) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

    # Create a separate data frame with the background colors
  bg_colors <- data.frame(
    from = factor(c("low-trust", "medium-trust", "high-trust"), levels=c("low-trust","medium-trust","high-trust"))
  )

  # plotting code...
  ggplot() +
    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
    ylim(c(0,1)) +
    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                    name = "From state") +  # Changed legend title for 'fill' here
    scale_color_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                       labels = c("low-trust", "medium-trust", "high-trust"),
                       name = "State transitioned to") +
    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(1, 'lines'),
          legend.spacing.x = unit(0.1, 'in'),
          legend.title = element_text(size = 14),
          legend.margin = margin(t = 0.2, b = 0, unit = 'cm'),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
          strip.text = element_text(size = 10),
          legend.box = "vertical" # Arrange legends vertically
    ) +
    guides(fill = guide_legend(order = 1, title.position = "left", title.hjust = 0.3),
           color = guide_legend(order = 2, title.position = "left", title.hjust = 0.3))
}

```

```{r, include=F}

plotInvTran <- plot_HMM_transitions(3, pars_inv) 
print(plotInvTran)

```

\begin{landscape}
```{r HMMPanels, fig.cap= "\\small{A: Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee decides how much to send back to the investor. Immediately before making this decision, participant are shown a two dimensional grid (Panel C) where they need to report their reaction to the investment. Those in the intervention condition were asked questions about their emotional state and those in the control condition were asked to report features of the investment unrelated to their emotion. Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees. From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel C. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel D. Each plot in Panel D represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.}",echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}

library(magick)

# Read in the images
image1 <- image_read("figures/trust_game.png")
image2 <- image_read("figures/HMMinvPolicy2.png")
image3 <- image_read("figures/emotion_grid.png")
image4 <- image_read("figures/HMMinvTrans2.png")

# Determine the target width for all images, for example, the width of the widest image
target_width <- max(image_info(image1)$width, image_info(image2)$width, 
                    image_info(image3)$width, image_info(image4)$width)

target_height <- max(image_info(image1)$height, image_info(image2)$height, 
                    image_info(image3)$height, image_info(image4)$height)

# Scale images to have the same width
image1 <- image_scale(image1, paste0(target_width, "x", target_height))
image2 <- image_scale(image2, paste0(target_width, "x", target_height))
image3 <- image_scale(image3, paste0(target_width, "x", target_height))
image4 <- image_scale(image4, paste0(target_width, "x", target_height))

# Annotate each image with a label
image1 <- image_annotate(image1, text = "A", location = "+200+10", size = 90, color = "black", gravity = "northwest")
image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")

# Assuming a 300 DPI resolution for good print quality
dpi <- 300

# Spacing of 0.1 inches converted to pixels
spacing <- 0.05 * dpi

# Calculating the width and height for each image
width_per_image <- ((12 - 0.1) / 2) * dpi
height_per_image <- ((7.4 - 0.1) / 2) * dpi
# The geometry string for each image, including spacing (assuming no borders)
geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)

# Now create the montage with the specified geometry
combined_image <- image_montage(
  c(image1, image2, image3, image4), 
  tile = "2x2",
  geometry = geometry_string
)

# Display the combined image
plot(combined_image)
```
\end{landscape}

In this game, the computerised agent was programmed to act according to a tit-for-tat strategy [@axelrod_evolution_1981], starting with a cooperative action and then mirroring what the other player chose in the preceding round. On round 4, the computerised agent was pre-programmed to choose the defect action, regardless of the participant's preceding action.

## Intervention

The intervention was built on Dialectical Behavior Therapy (DBT) skills training, asking patients to reflect on the consequences of actions taken in emotional states [@linehan_dbt_2015]. Specifically, participants were presented with a hypothetical situation in which they receive a low investment and asked to indicate how they would respond. They were then presented with an educational slide inviting them to consider that the ultimate aim in the game is to maximise their total reward and to reflect on whether punishing the investor for the low investment is beneficial to achieving that aim. Participants were told that punishment can create a negative feedback loop where the other player might trust them even less. An alternative action was suggested, whereby players would respond kindly to such a transgression in the hope of gaining trust from the investor. Participants were then asked whether the information just received would change their behavior in such a hypothetical situation and to justify their answer. Full details on the intervention are provided in the Supplementary Information.

In order to separate any general practice effects from the effect of the intervention, we added a control condition in which participants  were asked to solve five anagrams ("listen", "triangle", "deductions", "players", "care"). They provided their answers in a free-form text box. The time given to solve the anagrams was the same as that given to respond to questions in the intervention manipulation.

## Procedure

At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round Repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to evaluate their emotion (intervention condition) or the investment (control condition). Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived the investor to be (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the Repeated Prisoner's Dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities, emotion regulation, and BPD traits (see the supplement for details). They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

## Statistical analysis

To explore whether participants behaved differently in the RTG after the intervention compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model as described below:

\[
\begin{split}
R_{ij} = & \beta_0 + \beta_1 \text{ (Condition)}_i + \beta_2 \text{ (Game)}_i + \beta_3 \text{ (Investment)}_i + \\
& \beta_4 \text{ (Condition × Game)}_i + \beta_5 \text{ (Condition × Investment)}_i + \beta_6 \text{ (Game × Investment)}_i +  \\ 
& \beta_7 \text{ (Condition × Game × Investment)}_i +b_{0j} + b_{1j} \text{ (Game)}_i + \epsilon_{ij}
\end{split}
\]

where:

- \( R_{ij} \): percentage of tripled investment returned to investor for participant \( j \) in observation \( i \)
- \( \beta_0 \): intercept
- \( \beta_1 \): effect of Condition (intervention vs. control)
- \( \beta_2 \): effect of Game (RTG game pre vs. post-intervention)
- \( \beta_3 \): effect of Investment
- \( \beta_4 \): interaction effect between Condition and Game
- \( \beta_5 \): interaction effect between Condition and Investment
- \( \beta_6 \): interaction effect between Game and Investment
- \( \beta_7 \): three-way interaction effect between Condition, Game and Investment
- \( b_{0j} \): participant-wise random intercept for participant \( j \)
- \( b_{1j} \): participant-wise random slope for Game for participant \( j \)
- \( \epsilon_{ij} \): error term for participant \( j \) in observation \( i \)

The model was estimated using the `afex` package [@singmann_afex_2022] in R [@R_package]. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek2017balancing]. A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

To model participants' returns in the RTG across games and conditions, we fit various hidden Markov models [@visser2022hidden] to participants' returns using the depmixS4 package [@visser_depmixs4_2021] for R. The transition between latent states is assumed to depend on the investment received and a dummy variable to characterise the group that the participant belongs to. Details on how the models are constructed can be found in the supplement. We fit models with different numbers of hidden states, and use the Bayesian Information Criterion [@schwarz_estimating_1978-1] to select the best model.

# Behavioral results

```{r dataClean, include=FALSE}
# Filter only trust rounds and create % returns and investments

avg_ret_df <- d_finished %>%
  dplyr::select(playerId, condition.f,roundType,investment,returns,roundNum,gameNum.f,BPD_trait,PBOR_score,DERS_score,RFQ_c,RFQ_u) %>% 
  mutate( condition.f = fct_recode(condition.f, "intervention" = "coaxing"))%>%
  filter(roundType=="trust",!is.na(roundNum)) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = scale(investment)) %>% 
  mutate(inv_pct = investment/20, ret_pct = returns/(3*investment), ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)))  %>% 
  # Creating Quartiles of % returns by participant over BOTH TRUST GAMES
  group_by(playerId) %>% 
  mutate(mean_pct_ret = mean(ret_pct), mean_abs_ret = mean(returns), mean_inv = mean(investment)) %>%
  ungroup() %>% 
  mutate(Quart_pct_ret = ntile(mean_pct_ret,4),
         Quart_abs_ret = ntile(mean_abs_ret,4),
         Quart_inv     = ntile(mean_inv, 4))

# %>%
#   filter(complete.cases(.))   # <- one player had NA in one round as a return, so we exclude him from analysis.

# Number of people left for analysis. 
nrow(avg_ret_df %>% filter(condition.f == "intervention"))/30
nrow(avg_ret_df %>% filter(condition.f == "control"))/30

# average investment for first 10 rounds pre intervention
avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct, na.rm=TRUE))
```

```{r gamesPlot, include=F, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}

# Create a data frame for the vertical linesr
vline_data <- data.frame(
  xintercept = c(12, 13),
  Defection_round = factor(c("Pre-manipulation", "Post-manipulation"), levels = c("Pre-manipulation", "Post-manipulation"))
)
# Plot
ggplot(avg_ret_df, aes(x=as.factor(roundNum), y=ret_pct, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", size = 0.6) +
  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),
                        name = "Defection round") +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  facet_wrap(~condition.f) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") +
  guides(color = guide_legend(title = "Trust Game Phase", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Defection round"))

```

<!-- Plotting individual differences   -->
```{r, include =F}
ret_post_int <- avg_ret_df %>% filter(roundNum == 13 , gameNum.f == "post")

library(ggplot2)
ggplot(ret_post_int, aes(x = "", y = ret_pct)) +
  geom_violin(fill="lightblue") +
  geom_jitter(width = 0.2, height = 0.02, size = 1, col = "black") +
  geom_hline(yintercept = 0.33, color = "red", linetype = "dashed") +
  theme_bw() +
  facet_wrap(~condition.f) +
  labs(x = NULL, y = "% Return", title = "")

```

```{r modAllReturns, include=FALSE, cache=TRUE}

mod_returns_pct <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), avg_ret_df, REML= TRUE, method="KR")
anova(mod_returns_pct)
summary(mod_returns_pct)

prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400))

cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```

```{r, include=F}

library(sjstats)
effectsize::eta_squared(mod_returns_pct)

```

```{r, include=F}
# Function to calculate y_position for geom_signif
calculate_y_position <- function(plot) {
  plot_data <- ggplot_build(plot)$data[[1]]
  max_y <- max(plot_data$ymax, na.rm = TRUE)
  y_position <- max_y + (max_y * 0.05) # Adding 5% of the maximum y-value as a margin
  return(y_position)
}
```

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Trust Game Phase") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.88, tip_length = 0.03, textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "*", y_position = 0.80, tip_length = 0.03, textsize = 3)  # Control Group

```

Average investments and returns prior to the "defection round" (Figure \ref{fig:TrustGameResults}.A) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the percentage returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post, `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:TrustGameResults}.C). This indicates the intervention was effective in increasing cooperative behavior. There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$inv_scaled)`,
such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition. There was also an Investment by Game interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled)`. 
Finally, we find a three way interaction between Game, Condition and Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

```{r modInvs, include=FALSE, cache=TRUE}
mod_invs <- mixed( investment ~ gameNum.f*condition.f  + (1 + gameNum.f| playerId), avg_ret_df, method="KR")
summary(mod_invs)
anova(mod_invs)

#cond_effect_inv <- pairs(emmeans::emmeans(mod_invs, c("condition.f"), by = "gameNum.f", pbkrtest.limit = 9570))

# pairs(emmeans::emmeans(mod_invs, c("condition.f"), by = "gameNum.f" ))
# 
# pairs(emmeans::emmeans(mod_invs,c("gameNum.f") , by = c("condition.f")))
# 
# pairs(emmeans::emmeans(mod_invs, c("gameNum.f")))
# pairs(emmeans::emmeans(mod_invs, c("condition.f")))

```

```{r, include=FALSE}

inv_all <- afex_plot(mod_invs, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), 
            legend_title = "Condition") + 
  theme_bw() + 
  ggtitle("All rounds investments") +
  labs(y = "HMM Investment", x = "Trust Game Phase") + 
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 1, tip_length = -0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 18, tip_length = 0.03,textsize = 3)   # Control Group

```

To explore the HMM investors' behavior across games and conditions, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition and Game and their interaction as fixed effects, and a similar random effects structure to the returns model. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:TrustGameResults}.D, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

```{r, include=FALSE, cache=TRUE}
# ALL ROUNDS BEFORE DEFECTION TRIAL
pre_int_data <- avg_ret_df %>% filter(roundType=="trust",!is.na(gameNum.f),(roundNum < 12 & gameNum.f =="pre")| (roundNum < 13 & gameNum.f =="post"))

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| playerId), pre_int_data, REML= TRUE, method="KR")
anova(mod_returns_pre)
summary(mod_returns_pre)
saveRDS(mod_returns_pre, "data/mod_returns_pre")

pairs_pre_int <- pairs(emmeans::emmeans(mod_returns_pre, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400, reverse=TRUE))
saveRDS(pairs_pre_int, "data/pairs_pre_int")

```

```{r, include=F}
pctRet_pre <- afex_plot(mod_returns_pre, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Trust Game Phase") + 
  ggtitle("Pre defection trials percentage returns")+
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") + 
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "***", y_position = 0.95, tip_length = 0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "n.s", y_position = 0.85, tip_length = 0.03,textsize = 3)   # Control Group

```

```{r, include =FALSE}

# Defection trial only 
coax_data <- avg_ret_df %>%  filter(roundType=="trust",!is.na(gameNum.f),(roundNum == "12" & gameNum.f =="pre")| (roundNum == "13" & gameNum.f =="post"))

coax_data %>% group_by(gameNum.f, condition.f) %>% summarise(avg_inv = mean(investment),avg_return = mean(returns))

# ALL ROUNDS AFTER DEFECTION TRIAL (INCLUSIVE)
post_coax_data <- avg_ret_df %>% filter(roundType=="trust",!is.na(gameNum.f),(roundNum >= 12 & gameNum.f =="pre")| (roundNum >= 13 & gameNum.f =="post"))

post_coax_data %>% group_by(gameNum.f, condition.f) %>% summarise(avg_inv = mean(investment),avg_return = mean(returns))

```

```{r, include=FALSE, cache=TRUE}

mod_postcoax_trust <- mixed(ret_pct ~ gameNum.f*condition.f*inv_scaled  + ( 1 + gameNum.f|playerId), post_coax_data)
anova(mod_postcoax_trust)

postcoax_bygame <- emmeans::emmeans(mod_postcoax_trust, ~gameNum.f, pbkrtest.limit = 9400)
postcoax_bycond <- emmeans::emmeans(mod_postcoax_trust, ~condition.f, pbkrtest.limit = 9400)

postcoax_interaction <- emmeans::emmeans(mod_postcoax_trust, c("gameNum.f"), by = "condition.f", pbkrtest.limit = 9400)
postcoax_interaction

pairs(emmeans::emmeans(mod_postcoax_trust, c("gameNum.f"), by = "condition.f"))
```

```{r, include=F}
saveRDS(summary(mod_postcoax_trust), "data/mod_returns_post")
```

```{r, include=FALSE}

pctRet_post <- afex_plot(mod_postcoax_trust, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin,
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre-manipulation", "Post-manipulation")), 
            legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Trust Game Phase") + 
  ggtitle("Post defection trials percentage returns") +
  theme(plot.title = element_text(size = 16)) +
  theme(legend.position = "bottom") +
  geom_signif(xmin = 0.8, xmax = 1.8, annotations = "n.s", y_position = 0.95, tip_length = 0.03,textsize = 3) +  # Intervention Group
  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 0.9, tip_length = 0.03,textsize = 3)   # Control Group

```

We next analysed returns separately for rounds prior to the pre-programmed low investment by the HMM (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed-effects model as before to returns before the defection largely replicates the results over all trials. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-defection trials between the first and second RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`. Full results are provided in the supplement.

Applying the same model to the returns after the defection (see the supplement for full results), we again find a significant interaction between Condition and Game. Participants in the control condition decreased their post-defection returns from the first to the second RTG, pre-post, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no significant change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`.

Taken together, we find that participants in the control condition sent lower percentage returns in the second game, despite the HMM investor sending on average higher investments in that game. Those in the intervention group returned higher percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention compared to the control condition were not purely driven by reciprocity towards higher investments, since we found a Condition by Game interaction whilst controlling for investment in the model, and a reduced effect of investment in the intervention condition. The intervention did not increase participants' returns after the defection by the other player. Instead, it countered the tendency shown by participants in the control condition to lower returns after the defection in the second game compared to the first game.

\newpage
\begin{landscape}
```{r TrustGameResults, fig.cap="\\small{Panel A: Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. We note a different reaction to the pre-programmed one-off low investment (vertical black lines) between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel. B:  Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's pre-programmed one-off low investment round (vertical grey lines) were similar before (round 12) and after (round 13) the intervention. The bottom panels show marginal means and distributions of either investments or percentage returns across participants by Game and Condition. Panel C shows that participants in the Intervention condition returned higher proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back lower returns. Investments by the HMM agent (Panel D) were higher in the second game compared to the first game across conditions.}", echo=FALSE, message=FALSE, warning=FALSE, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
library(magick)

# Read in the images
image5 <- image_read("figures/resultsTrust.png")
image6 <- image_read("figures/emotionReaction2.png")
violin_ret <- image_read("figures/violin_ret.png")
violin_inv <- image_read("figures/violin_inv.png")

image5 <- image_annotate(image5, text = "A", location = "+20+20", size = 70, color = "black", gravity = "northwest")
image6 <- image_annotate(image6, text = "B", location = "+20+0", size = 90, color = "black", gravity = "northwest")
violin_ret<- image_annotate(violin_ret, text = "C", location = "+0+20", size = 70, color = "black", gravity = "northwest")
violin_inv  <- image_annotate(violin_inv , text = "D", location = "+0+0", size = 70, color = "black", gravity = "northwest")

# Assuming a 300 DPI resolution for good print quality
dpi <- 300

# Spacing of 0.1 inches converted to pixels
spacing <- 0.05 * dpi

# Calculating the width and height for each image
width_per_image <- ((12 - 0.1) / 2) * dpi
height_per_image <- ((7.4 - 0.1) / 2) * dpi
# The geometry string for each image, including spacing (assuming no borders)
geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)

# Now create the montage with the specified geometry
combined_image2 <- image_montage(
  c(image5, image6, violin_ret, violin_inv), 
  tile = "2x2",
  geometry = geometry_string
)

# Display the combined image
plot(combined_image2)
```
\end{landscape}
\newpage

## Emotion self-reports

```{r, include=TRUE}

# Get Emotions on x axis per round (valence)
emo.x <- d_finished %>% 
  mutate( condition.f = fct_recode(condition.f, "intervention" = "coaxing")) %>% 
  dplyr::select(playerId,roundType,condition.f,contains("emotionGrid")) %>% 
  dplyr::filter(roundType=="trust") %>% dplyr::select (-contains(".y") ) %>% pivot_longer(cols=contains(".x"), names_to = c("gameNumber","roundNum"), names_pattern ="return_trust_emotionGrid_(.*)_(.*)[.]x", values_to = "emo_coord_x") %>%
  distinct()

#Get emotions on y axis (arousal)
emo.y <- d_finished %>%
  mutate( condition.f = fct_recode(condition.f, "intervention" = "coaxing")) %>%
  dplyr::select(playerId,roundType,condition.f,contains("emotionGrid")) %>% 
  dplyr::filter(roundType=="trust") %>% dplyr::select(-contains(".x") ) %>% pivot_longer(cols=contains(".y"), names_to = c("gameNumber","roundNum"), names_pattern ="return_trust_emotionGrid_(.*)_(.*)[.]y", values_to = "emo_coord_y") %>%
  distinct()

# put emotions in one dataframe
full_emo <- full_join(emo.x,emo.y, by=c("playerId", "gameNumber","roundNum","roundType", "condition.f"))  %>% 
  mutate(roundNum=as.numeric(as.character(roundNum))) %>% 
  mutate(gameNum.f = factor(gameNumber,labels = c("pre","post"),levels=c("1","2"))) %>% 
  dplyr::select(-c("roundType","gameNumber")) %>% 
  # group_by(condition.f) %>% 
  mutate(emo_scaled_x = scale(emo_coord_x), emo_scaled_y= -scale(emo_coord_y))    # <--- SCALING and reversing Y coordinates (0,0 is top left)

#merge with returns dataframe
full_dat <- full_join(avg_ret_df, full_emo, by = c("playerId","condition.f","gameNum.f","roundNum" )) %>% filter(complete.cases(.))
#nrow(full_dat)

datCoax <- full_dat %>% filter(condition.f=="intervention")
datCtrl <- full_dat %>% filter(condition.f=="control")

```

```{r emoPlotCoax, include=F, fig.cap="Self-reported emotion valence and arousal as well as investment z-scores for each round of the Repeated Trust Game averaged across participants in the intervention condition only. The participants' emotion reaction measured during the investor's defection round were similar before (round 12) and after (round 13) the intervention.",fig.align="center", fig.width=7, fig.height = 4}

#################### two panels for before and after defection 
library(ggplot2)

# Define the colors in alphabetical order
colors <- c("Arousal" = "forestgreen", "Investment" = "blue", "Valence" = "red")

# Create a new variable to categorize the rounds
datCoax$roundGroup <- ifelse(datCoax$roundNum <= 10, "Rounds 1-10", "Rounds 11-15")

# # Define the vertical line data
# vline_data <- data.frame(
#   xintercept = c(12, 13),
#   linetype = c("Pre-manipulation", "Post-manipulation"),  # Names corresponding to the existing scale
#   roundGroup = c("Rounds 11-15", "Rounds 11-15")
# )

p <- ggplot(datCoax[datCoax$gameNum.f == 'pre', ], aes(x=roundNum)) + 
  stat_summary(fun = "mean", geom = "line", group =1, aes(y=emo_scaled_x, color = "Valence", linetype = "a_Pre")) + 
  stat_summary(fun = "mean", geom = "line", group =1, aes(y=emo_scaled_y, color = "Arousal", linetype = "a_Pre")) + 
  stat_summary(fun = "mean", geom = "line", group =1, aes(y=inv_scaled, color = "Investment", linetype = "a_Pre")) +
  # Plot 'post'
  stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=emo_scaled_x, color = "Valence", linetype = "b_Post")) + 
  stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=emo_scaled_y, color = "Arousal", linetype = "b_Post")) + 
  stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=inv_scaled, color = "Investment", linetype = "b_Post")) +
  labs(x = "Round", y="Z-score") +
  geom_hline(yintercept = 0, linetype="dotted", color = "grey", linewidth=1.5) +
  scale_color_manual(name = "", values = colors, labels = c("Arousal", "Investment", "Emotion valence")) +
  scale_linetype_manual(values = c("a_Pre" = "dotted", "b_Post" = "solid"),
                        labels = c("Pre-manipulation", "Post-manipulation")) +
geom_vline(data = subset(datCoax, roundGroup == "Rounds 11-15" & roundNum == 12), aes(xintercept = roundNum), linetype = "dotted", color = "darkgrey", size = 0.6, show.legend = FALSE) +
  geom_vline(data = subset(datCoax, roundGroup == "Rounds 11-15" & roundNum == 13), aes(xintercept = roundNum), linetype = "solid", color = "darkgrey", size = 0.6, show.legend = FALSE) +
  scale_x_continuous(breaks = 1:max(datCoax$roundNum)) +
  facet_wrap(~roundGroup, scales = "free_x") +
  theme_bw() + 
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Trust Game Phase"))

# Print the plot
print(p)

```

```{r, include=FALSE, cache=TRUE}

mod_emo_x <- mixed( emo_scaled_x ~ gameNum.f*scale(investment) + (1 + gameNum.f| playerId), data=datCoax, method="KR")
summary(mod_emo_x)
anova(mod_emo_x)

modVal <- emmeans::emmeans(mod_emo_x, ~gameNum.f)
pairs(modVal)

mod_emo_y <- mixed( emo_scaled_y ~ gameNum.f*scale(investment) + (1 + gameNum.f| playerId), datCoax )
summary(mod_emo_y)
anova(mod_emo_y)

modArousal <- emmeans::emmeans(mod_emo_y, ~gameNum.f)
pairs(modArousal)

# afex_plot(mod_emo_x, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
#             data_geom = geom_violin, 
#             data_arg = list(width = 0.5))
#

# library(GGally)
# ggpairs(datCoax %>% dplyr::select(emo_scaled_x,emo_scaled_y,inv_scaled))

cor(datCoax %>% dplyr::select(emo_scaled_x,emo_scaled_y,investment))
cor.test(datCoax$emo_scaled_x, datCoax$emo_scaled_y)
```

To assess the impact of the intervention on the participant's emotional reactions, we used linear mixed-effects models (one for valence, and one for arousal) with fixed effects for Game (pre vs. post intervention) and Investment, as well as interaction between Investment and Game, with participant-wide random intercepts and random slopes for Game. This showed that higher investments were associated with more positive emotions, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$scaleinvestment)`, 
and higher arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$scaleinvestment)`. 
In addition, the positiveness of emotion declined between the two games,
`r papaja::apa_print(mod_emo_x)$full_result$gameNum_f`
, as did arousal,
`r papaja::apa_print(mod_emo_y)$full_result$gameNum_f`. 
There was no indication that the effect of the investment on either aspect of emotion was affected by the intervention, as there was no interaction between Investment and Game on valence, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_x)$full_result$gameNum_f_scaleinvestment)`, or arousal, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_emo_y)$full_result$gameNum_f_scaleinvestment)`. 
This indicates that participants in the intervention condition returned higher amounts post-intervention, despite their emotional reaction to investments remaining largely the same (Figure \ref{fig:TrustGameResults}.B).

## Evaluation of the investor

```{r, include=FALSE}

# Pivot longer ratings on each attribute 
df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative")) %>%  
  pivot_longer(cols=contains("cooperative"), names_to = c("gameNumber"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()

df_selfish <- d_finished %>% dplyr::select(playerId,condition.f,contains("selfish")) %>% 
  pivot_longer(cols=contains("selfish"), names_to = c("gameNumber"), names_pattern ="rating_selfish_(.*)", values_to = "rating_selfish") %>% distinct

df_trustworthy <- d_finished %>% dplyr::select(playerId,condition.f,contains("trustworthy")) %>% 
  pivot_longer(cols=contains("trustworthy"), names_to = c("gameNumber"), names_pattern ="rating_trustworthy_(.*)", values_to = "rating_trustworthy") %>% distinct()

df_friendly  <-  d_finished %>% dplyr::select(playerId,condition.f,contains("friendly")) %>% 
  pivot_longer(cols=contains("friendly"), names_to = c("gameNumber"), names_pattern ="rating_friendly_(.*)", values_to = "rating_friendly") %>%  distinct()

#merge all data frames together
datRatings <- list(df_coop, df_selfish,df_trustworthy, df_friendly) %>% 
              reduce(full_join, by=c('playerId','gameNumber','condition.f') ) %>% 
              mutate(gameNum.f = factor(gameNumber,labels = c("pre","post"),levels=c("1","2")))

# Group means for each rating 
mu <- datRatings %>% group_by(gameNum.f, condition.f) %>% summarise(mean_coop = mean(rating_coop),
                                                  mean_selfish = mean(rating_selfish),
                                                  mean_trustworthy = mean(rating_trustworthy),
                                                  mean_friendly = mean(rating_friendly)
                                                  )
```

```{r, include=FALSE}

# Cooperative
coop_mod <- mixed(rating_coop ~ gameNum.f*condition.f + (1 | playerId), datRatings)
summary(coop_mod)

pairs(emmeans::emmeans(coop_mod, c("gameNum.f"), by = "condition.f"))

#Selfish
self_mod <- mixed(rating_selfish ~ gameNum.f*condition.f + (1 | playerId), datRatings)
summary(self_mod)
pairs(emmeans::emmeans(self_mod, c("gameNum.f"), by = "condition.f"))

#Trustworthy 
trust_mod <- mixed(rating_trustworthy ~ gameNum.f*condition.f + (1 | playerId), datRatings)
summary(trust_mod)
pairs(emmeans::emmeans(trust_mod, c("gameNum.f"), by = "condition.f"))

#Friendly
friend_mod <- mixed(rating_friendly ~ gameNum.f*condition.f + (1 | playerId), datRatings)
summary(friend_mod)
pairs(emmeans::emmeans(friend_mod, c("gameNum.f"), by = "condition.f"))

```

For the Investor evaluation, we estimate a mixed-effects model for participants ratings with Game and Condition as fixed effects and participant-wise random intercepts as random effects. Participants rated the HMM investor in the second game as less cooperative (`r apa_print(pairs(emmeans::emmeans(coop_mod, c("gameNum.f"))))$full_result`), less trustworthy (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("gameNum.f"))))$full_result`), less friendly (`r apa_print(pairs(emmeans::emmeans(friend_mod, c("gameNum.f"))))$full_result`) and more selfish (`r apa_print(pairs(emmeans::emmeans(self_mod, c("gameNum.f"))))$full_result`), than the HMM investor in the first game. Participants in the intervention condition rated players higher than those in the control condition on cooperativeness (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("condition.f"))))$full_result`) and lower on selfishness (`r apa_print(pairs(emmeans::emmeans(self_mod, c("condition.f"))))$full_result`). There was no evidence for an interaction effect between Condition and Game on any of the attributes.

When asked during debrief whether they thought the investors they faced were Human or not, $40$% of participants thought they were either facing a human or were not sure of the nature of the co-player. Many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.

## Transfer to the Repeated Prisoner's Dilemma game

```{r, include=FALSE}

PD_data <- d_finished %>%
  mutate( condition.f = fct_recode(condition.f, "intervention" = "coaxing")) %>%
  dplyr::select(playerId, condition.f,roundType,roundNum, PDoption,PDchoice, AiChoicePD,gameNum.f, PBOR_score) %>% 
                          filter(roundType == "PD",!is.na(roundNum)) %>% 
                          dplyr::mutate(coop = ifelse(PDchoice=="cooperate",1,0),phase = ifelse(roundNum <= 4, 0,1))

```

```{r PDCoop, include=FALSE, fig.cap="Mean and standard error of the rate at which the cooperative action was chosen by the participants for each round of the Repeated Prisoner's Dilemma game. Round 4 is where we programmed the Tit-for-Tat agent to defect, which explains the lower cooperation rate we see in round 5 onwards",fig.align="center", fig.width=6, fig.height =3}
# plot of cooperation rates per round
ggplot(PD_data,aes(x=roundNum,group = 1)) + 
  stat_summary(fun = "mean", geom = "line", aes(y=coop), color ="red") + 
  stat_summary(fun.data = "mean_se", geom = "errorbar",alpha =0.3, width = 0.2, aes(y=coop) ) + 
  facet_grid(~ condition.f) +
  xlab("Round") + 
  ylab("Cooperation rate (%)") +
  scale_x_continuous(breaks = round(seq(1,10, by = 1),1)) + 
  theme_bw()

```

```{r, include=FALSE}

ipd_mod <- mixed(coop ~ condition.f * phase + (1 |playerId), data= PD_data, method = "LRT", family= "binomial") 
summary(ipd_mod)

ipd_mod_r5 <- glm(coop ~ condition.f, data= PD_data%>% filter(roundNum >= 5)) 
summary(ipd_mod_r5)

```

We next asked whether the intervention had any discernible effect on participants' behavior in a different, Repeated Prisoner's Dilemma game. Predicting the probability of a cooperative action with a logistic mixed-effects regression model, with Condition and Phase (before or after defection trial) as fixed effects and a random intercept for participants, showed a decline in cooperation after defection by the other player, `r papaja::apa_print(ipd_mod)$full_result$phase`, but no evidence for an overall different cooperation rate in the intervention condition compared to the control condition, `r papaja::apa_print(ipd_mod)$full_result$condition_f`, or a different response to defection between the conditions, `r papaja::apa_print(ipd_mod)$full_result$condition_f_phase`. As such, there is no evidence that the intervention affected behavior in this game.

```{r, include=FALSE}
########### HOW DO PBOR SCORES affect returns and intervention effect ?########

avg_ret_df <- avg_ret_df %>% mutate(PBOR_scaled = scale(PBOR_score), DERS_scaled = scale(DERS_score), RFQc_scaled = scale(RFQ_c))

# these are repeated measures here, so not independent. need to used mixed effects model.
mixed_PBOR_pct <- afex::mixed(ret_pct ~ PBOR_scaled*gameNum.f*condition.f + investment + (1 | playerId ), data= avg_ret_df)
summary(mixed_PBOR_pct)

# At average value of PBOR score  
emmeans(mixed_PBOR_pct, pairwise ~ PBOR_scaled * condition.f )

## Looking at difference between average returns across rounds
#scoreDAT <- avg_ret_df %>% dplyr::select(playerId,diff_pctRet, PBOR_scaled, DERS_scaled, RFQ_c, condition.f) %>% unique()

```

```{r, include=FALSE}
# Install the 'psych' package
if (!requireNamespace("psych", quietly = TRUE)) {
  install.packages("psych")
}

# Load the 'psych' package
library(psych)

questionnaires <- avg_ret_df %>% group_by(playerId) %>% dplyr::select(PBOR_score,DERS_score,RFQ_c) %>% unique() %>% ungroup()
# Perform EFA
efa_result <- fa(questionnaires %>% dplyr::select(PBOR_score,DERS_score,RFQ_c), nfactors = 2, fm = "minres", rotate = "varimax")
# Print the EFA result
print(efa_result)

```

```{r, include=FALSE}
turing <- read_csv("data/turing.csv")
colnames(turing) <- c("responses", "category")

turing  %>%
    group_by(category) %>%
    summarize(count = n()) %>%
    mutate(percentage = count / sum(count) * 100) %>%
    dplyr::select(category, percentage)

```

# HMM analysis of participant returns

```{r, include=FALSE, echo=FALSE, warning=F}

# Removing player whose return was not properly recorded by data (NA in one of the rounds) so as to run depmixS4
id_with_na <-  as.character(avg_ret_df[is.na(avg_ret_df$returns),"playerId"])

# Create return percentage bins Bins
avg_ret_df <- avg_ret_df %>%
  filter(playerId != id_with_na) %>% 
  mutate(game_id = rep(1:636, each=15)) %>%
  group_by(game_id) %>%
  mutate(next_investment = lead(investment, default=0),
         investment_bin = cut(investment, breaks = c(-1,2.5,7.5,12.5,17.5,21)),  
         return_pct_bin = cut(ret_pct_0, breaks = c(-.1,.16,.33,.50,.66,0.84,1.1)),  # 0, 1/6, 2/6, 3/6,...
         intervention_ctrst = factor(ifelse(gameNum.f=="pre", 0,ifelse(condition.f=="control",1,2))),
         ctrl_only_ctrst = factor(ifelse(gameNum.f=="post" & condition.f=="control", 1,0)),
         coax_only_ctrst = factor(ifelse(gameNum.f=="post" & condition.f=="intervention", 1,0)),
         pre_post_ctrst = factor(ifelse(gameNum.f=="post", 1,0)))  %>% 
  ungroup()
# We put factor as otherwise R thinks that it's a linear variable. We want the first level to be reference group, 0. Default uses that.

priordat <- avg_ret_df %>% filter(roundNum==1)

############### HMM support functions   #########################

# order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
label_switch <- function(mod,labels) {
  # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
  if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
  n_states <- mod@nstates
  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
    stop("labels needs to be a vector of unique integers between 1 and", n_states)
  }
  inv_labels <- sapply(1:n_states,function(x) which(labels == x))
  tmp <- mod
  # relabel prior
  ppars <- getpars(mod@prior)
  fpars <- getpars(mod@prior,which="fixed")
  out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
  out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
  if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
  # relabel transition
  for(i in 1:n_states) {
    ppars <- getpars(mod@transition[[inv_labels[i]]])
    fpars <- getpars(mod@transition[[inv_labels[i]]],which="fixed")
    out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
    out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
    if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
    #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
  }
  # relabel response
  for(i in 1:n_states) {
    out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
    out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
  }
  tmp <- setpars(tmp,out_fixed,which="fixed")
  tmp <- setpars(tmp,out_pars)
  if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
  return(tmp)
}

# ORDERING STATES
order_mod_gauss <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation 
    sum[i] = mod@response[[i]][[1]]@parameters$mu
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

# ordering states for a truncated discretised Gaussian
order_mod_truncdiscgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- truncdiscgaus(seq(0, 20), pstart = tpars, min=-0.5, max=20.5)
    sum[i] <- sum(seq(0,20)*dens(dmod))
  }
  cat("Expected values for states: ",sum, "\n")
  # reordering the states
  mod <- label_switch(mod, rank(sum, ties.method = "first"))
  return(mod)
}

# ordering states for a variable response truncated discretised Gaussian
order_mod_vtdgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- vtdgaus(seq(0, 1, length=61), pstart = tpars, yield=rep(60, 61))
    sum[i] <- sum(seq(0, 1, length=61)*dens(dmod))
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum, ties.method = "first"))
  return(mod)
}

```

```{r,include=F}
# define a response class which only contains the standard slots, no additional slots
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

# define a generic for the method defining the response class

setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))

# define the method that creates the response class

setMethod("discgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu = 10, sigma = 3)
            }
            mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
            mod
          }
)

setMethod("show","discgaus",
          function(object) {
            cat("Gaussian with discrete support\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

setMethod("dens","discgaus",
          function(object,log=FALSE) {
            p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

setMethod("setpars","discgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values)!=npar) stop("length of 'values' must be",npar)
            # determine whether parameters or fixed constraints are being set
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

setMethod("getpars","discgaus",
          function(object,which="pars",...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

setMethod("fit","discgaus",
          function(object,w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object,c(pars[1], exp(pars[2])))
            object
          }
)

setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))

setMethod("dens","truncdiscgaus",
          function(object,log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            prec <- pnorm(object@max, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@min, mean = object@parameters$mu, sd = object@parameters$sigma)
            if(prec < 1e-12) {
              # return uniform probabilities as precision is likely off
              p <- rep(1/(length(breaks)-1), length(object@y))
              if(log) return(log(p)) else return(p)
            } else {
              p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/sum(p)
              if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
            }
          }
)

setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))

setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)

# define a response class which only contains the standard slots, no additional slots
setClass("vtdgaus", contains="response", slots=c(yield="numeric"))

# define a generic for the method defining the response class

setGeneric("vtdgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("vtdgaus"))

# define the method that creates the response class

setMethod("vtdgaus", 
          signature(y="ANY"), 
          function(y, yield, pstart=NULL,fixed=NULL, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu=.5,sigma=1)
            }
            mod <- new("vtdgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, yield=yield)
            mod
          }
)

setMethod("show","vtdgaus",
  function(object) {
    cat("Gaussian with variable discrete support for percentage responses\n")
    cat("Parameters: \n")
    cat("mu: ", object@parameters$mu, "\n")
    cat("sigma: ", object@parameters$sigma, "\n")
  }
)

setMethod("dens","vtdgaus",
  function(object,log=FALSE) {
      # determine cuts for pnorm based on 
      #p <- pnorm(sapply(object@y + .5*(1/object@yield), function(x) min(x,1.001)), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(sapply(object@y - .5*(1/object@yield),function(x) max(x,-0.001)), mean = object@parameters$mu, sd = object@parameters$sigma)
      #norm <- (pnorm(1 + .001, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(-0.001, mean = object@parameters$mu, sd = object@parameters$sigma))
      #p <- p/norm

      # alternatively, normalize factor based on size of cuts

      prec <- pnorm(1 + .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma)
      # check for issues with precision, using highest yield
      if(prec < 1e-12) {
        p <- 1/(object@yield + 1)
      } else {
        p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
        norm <- pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
        p <- p/norm
      }
      # probability when yield == 0 is always 1 
      p[object@yield == 0] <- 1
      if(log) return(log(p)) else return(p)
    }
)

setMethod("setpars","vtdgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values)!=npar) stop("length of 'values' must be",npar)
            # determine whether parameters or fixed constraints are being set
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

setMethod("getpars","vtdgaus",
          function(object,which="pars",...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

setMethod("fit","vtdgaus",
          function(object,w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object,c(pars[1], exp(pars[2])))
            object
          }
)
```

```{r, include=F}
priordat <- avg_ret_df %>% filter(roundNum==1)
n_state_min <- 2
n_state_max <- 7

trust_simple <- trust_ctrl <- trust_coax <- trust_fullCtrst <- trust_prePost <- list()

# HMMs based on bins
for(i in n_state_min:n_state_max) {

  trust_simple[[i]] <- depmix(return_pct_bin ~ 1, data = avg_ret_df, nstates = i, transition = ~ next_investment, prior = ~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes = rep(15,636))

  trust_ctrl[[i]] <- depmix(return_pct_bin ~ 1, data = avg_ret_df, nstates = i, transition = ~ next_investment*ctrl_only_ctrst, prior = ~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes = rep(15,636))

  trust_coax[[i]] <- depmix(return_pct_bin ~ 1, data = avg_ret_df, nstates = i, transition = ~ next_investment*coax_only_ctrst, prior = ~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes = rep(15,636))

  trust_fullCtrst[[i]] <- depmix(return_pct_bin ~ 1, data = avg_ret_df, nstates = i, transition = ~ next_investment*intervention_ctrst, prior = ~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes = rep(15,636))

  trust_prePost[[i]] <- depmix(return_pct_bin ~ 1, data = avg_ret_df, nstates = i, transition = ~ next_investment*pre_post_ctrst, prior = ~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes = rep(15,636))

}

# vtdgaus no contrasts
simple_HMMs <- ctrl_HMMs <- coax_HMMs <- fullCtrst_HMMs <- prePost_HMMs <- list()

for(i in n_state_min:n_state_max) {

  # rModels <- rep(list(list(vtdgaus(y=avg_ret_df$ret_pct_0,yield=3*avg_ret_df$investment))),each=i)

  ###########################################################
  mus <- (1:i)*(1/(i+1))
  sigmas <- rep(1/(i+1), i)
  rModels <- list()
  for(s in 1:i) {
    rModels[[s]] <- list(vtdgaus(y=avg_ret_df$ret_pct_0,yield=3*avg_ret_df$investment, pstart=c(mus[s],sigmas[s])))
  }
  #########################################################

  simple_HMMs[[i]] <- makeDepmix(response=rModels,transition=trust_simple[[i]]@transition,prior=trust_simple[[i]]@prior, ntimes = rep(15,636), homogeneous=FALSE)

  ctrl_HMMs[[i]] <- makeDepmix(response=rModels,transition=trust_ctrl[[i]]@transition,prior=trust_ctrl[[i]]@prior, ntimes = rep(15,636), homogeneous=FALSE)

  coax_HMMs[[i]] <- makeDepmix(response=rModels,transition=trust_coax[[i]]@transition,prior=trust_coax[[i]]@prior, ntimes = rep(15,636), homogeneous=FALSE)

  fullCtrst_HMMs[[i]] <- makeDepmix(response=rModels,transition=trust_fullCtrst[[i]]@transition,prior=trust_fullCtrst[[i]]@prior, ntimes = rep(15,636), homogeneous=FALSE)

  prePost_HMMs[[i]] <- makeDepmix(response=rModels,transition=trust_prePost[[i]]@transition,prior=trust_prePost[[i]]@prior, ntimes = rep(15,636), homogeneous=FALSE)

}

```

```{r, include=FALSE}

# fittedSimple <- fittedCoax  <- fittedCtrl  <- fittedFullCtrst  <- fittedPrePost  <- list()
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
# 
#   #fittedSimple[[i]] <- multistart(simple_HMMs[[i]],nstart=20)
#   fittedSimple[[i]] <- fit(simple_HMMs[[i]], emcontrol=em.control(random.start = FALSE))
# }
# save(fittedSimple, file="data/fittedSimple.RData")
# ##############################
# 
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
#   #fittedCtrl[[i]] <- multistart(ctrl_HMMs[[i]],nstart=20)
#   fittedCtrl[[i]] <- fit(ctrl_HMMs[[i]], emcontrol=em.control(random.start = FALSE))
# }
# save(fittedCtrl, file="data/fittedCtrl.RData")
# ##############################
# 
# 
# set.seed(20230918)
# for(i in n_state_min:n_state_max) {
#   # fittedCoax[[i]] <- multistart(coax_HMMs[[i]],nstart=20)
#   fittedCoax[[i]] <- fit(coax_HMMs[[i]], emcontrol=em.control(random.start = FALSE))
# }
# save(fittedCoax, file="data/fittedCoax.RData")
# ##############################
# 
# set.seed(20221028)
# for(i in n_state_min:n_state_max) {
#   #fittedPrePost[[i]] <- multistart(prePost_HMMs[[i]],nstart=20)
#   fittedPrePost[[i]] <- fit(prePost_HMMs[[i]], emcontrol=em.control(random.start = FALSE))
# }
# save(fittedPrePost, file="data/fittedPrePost.RData")
# #############################
# 
# set.seed(20230922)
# for(i in n_state_min:n_state_max) {
#   #fittedFullCtrst[[i]] <- multistart(fullCtrst_HMMs[[i]],nstart=20)
#   fittedFullCtrst[[i]] <- fit(fullCtrst_HMMs[[i]], emcontrol=em.control(random.start = FALSE))
# }
# save(fittedFullCtrst, file="data/fittedFullCtrst.RData")
# ###############################

```

```{r, include=FALSE}

load("data/fittedSimple.RData")
load("data/fittedCtrl.RData")
load("data/fittedCoax.RData")
load("data/fittedPrePost.RData")
load("data/fittedFullCtrst.RData")

```

```{r, include=F}

# Fitting "Ctrl" models with initial parameters from the fitted "Simple" model.

# newCtrl <- fittedCtrl
# for (ns in n_state_min : n_state_max){
# 
#   fmod <- fittedSimple[[ns]]
#   origin_pars <- getpars(fmod)
#   numCovIn <- 2
#   numCovTr <- 2
# 
#   # create a new vector which has the right number of pars for the initial state probabilities
#   new_pars <- c(origin_pars[1:(ns*numCovIn)])
#   # Build a new vector for transition probabilities, leaving space for new contrasts. "from" is state we are transitioning from
#   for(from in 1:ns) {
#   mat <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)],ncol=numCovTr)
#   #print(mat)
#   new_mat <- mat[,c(1:2,2,2)]
#   new_pars <- c(new_pars,as.numeric(new_mat))
#   }
#   #Concatenate the new vectors with the old vector of  "response parameters"
#   new_pars <- c(new_pars,origin_pars[(ns*numCovIn + 1 + (ns - 1)*numCovTr*ns + numCovTr*ns):length(origin_pars)])
# 
#   # Set initial parameters of more complex model as the matrix of parameters we created from the simpler model and fit it.
#   newCtrl[[ns]] <- setpars(newCtrl[[ns]],new_pars)
#   newCtrl[[ns]] <- fit(newCtrl[[ns]], emcontrol= em.control(maxit = 1000,random.start = FALSE))
# }
# 
# save(newCtrl, file="data/newCtrl.RData")

```

```{r, include=F}

# Fitting "Coax" models with initial parameters from the fitted "Simple" model.
# newCoax <- fittedCoax
# for (ns in n_state_min : n_state_max){
# 
#   fmod <- fittedSimple[[ns]]
#   origin_pars <- getpars(fmod)
#   numCovIn <- 2
#   numCovTr <- 2
# 
#   # create a new vector which has the right number of pars for the initial state probabilities
#   new_pars <- c(origin_pars[1:(ns*numCovIn)])
#   # Build a new vector for transition probabilities, leaving space for new contrasts. "from" is state we are transitioning from
#   for(from in 1:ns) {
#   mat <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)],ncol=numCovTr)
#   #print(mat)
#   new_mat <- mat[,c(1:2,2,2)]
#   new_pars <- c(new_pars,as.numeric(new_mat))
#   }
#   #Concatenate the new vectors with the old vector of  "response parameters"
#   new_pars <- c(new_pars,origin_pars[(ns*numCovIn + 1 + (ns - 1)*numCovTr*ns + numCovTr*ns):length(origin_pars)])
# 
#   # Set initial parameters of more complex model as the matrix of parameters we created from the simpler model and fit it.
#   newCoax[[ns]] <- setpars(newCoax[[ns]],new_pars)
#   newCoax[[ns]] <- fit(newCoax[[ns]], emcontrol= em.control(maxit = 2000,random.start = FALSE))
# }
# 
# save(newCoax, file="data/newCoax.RData")

```

```{r, include=F}

# Fitting "Prepost" models with initial parameters from the fitted "Simple" model.

# newPrePost <- fittedPrePost
# for (ns in n_state_min : n_state_max){
# 
#   fmod <- fittedSimple[[ns]]
#   origin_pars <- getpars(fmod)
#   numCovIn <- 2
#   numCovTr <- 2
# 
#   # create a new vector which has the right number of pars for the initial state probabilities
#   new_pars <- c(origin_pars[1:(ns*numCovIn)])
#   # Build a new vector for transition probabilities, leaving space for new contrasts. "from" is state we are transitioning from
#   for(from in 1:ns) {
#   mat <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)],ncol=numCovTr)
#   #print(mat)
#   new_mat <- mat[,c(1:2,2,2)]
#   new_pars <- c(new_pars,as.numeric(new_mat))
#   }
#   #Concatenate the new vectors with the old vector of  "response parameters"
#   new_pars <- c(new_pars,origin_pars[(ns*numCovIn + 1 + (ns - 1)*numCovTr*ns + numCovTr*ns):length(origin_pars)])
# 
#   # Set initial parameters of more complex model as the matrix of parameters we created from the simpler model and fit it.
#   newPrePost[[ns]] <- setpars(newPrePost[[ns]],new_pars)
#   newPrePost[[ns]] <- fit(newPrePost[[ns]], emcontrol= em.control(maxit = 1000,random.start = FALSE))
# }
# 
# save(newPrePost, file="data/newPrePost.RData")

```

```{r, include=FALSE}

# For initialising newFULL, take the model with the highest likelihood in each num states:

```

```{r, include=FALSE}

load("data/newCtrl.RData")
load("data/newCoax.RData")
load("data/newPrePost.RData")

```

```{r, include=F}

bestCtrl <- NULL
# Loop through each index and select the model with the higher log-likelihood
for(i in n_state_min : n_state_max) {
  if(logLik(fittedCtrl[[i]]) > logLik(newCtrl[[i]])) {
    bestCtrl[[i]] <- fittedCtrl[[i]]
  } else {
    bestCtrl[[i]] <- newCtrl[[i]]
  }
}

bestCoax <- NULL
# Loop through each index and select the model with the higher log-likelihood
for(i in n_state_min : n_state_max) {
  if(logLik(fittedCoax[[i]]) > logLik(newCoax[[i]])) {
    bestCoax[[i]] <- fittedCoax[[i]]
  } else {
    bestCoax[[i]] <- newCoax[[i]]
  }
}

bestPrePost <- NULL
# Loop through each index and select the model with the higher log-likelihood
for(i in n_state_min : n_state_max) {
  if(logLik(fittedPrePost[[i]]) > logLik(newPrePost[[i]])) {
    bestPrePost[[i]] <- fittedPrePost[[i]]
  } else {
    bestPrePost[[i]] <- newPrePost[[i]]
  }
}

```

```{r, include=FALSE}

# # Here we use parameters of BEST PrePost model as initial values of the FullContrast model to check whether we can approach true maximal likelihood.

# newFull <- fittedFullCtrst
# for (ns in n_state_min :n_state_max){
# 
#   # Use the higher likelihood prepost model parameters as starting values to fit the newFull model.
#   fmod <- bestPrePost[[ns]]
#   origin_pars <- getpars(fmod)
#   numCovTr <- 4
#   numCovIn <- 2
# 
#   # create a new vector which has the right number of pars for the full contrast model.
#   new_pars <- c(origin_pars[1:(ns*2)])
#   # from is state we are transitioning from
#   for(from in 1:ns) {
#   mat <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)],ncol=numCovTr)
#   #print(mat)
#   new_mat <- mat[,c(1:2,3,3,4,4)]
#   new_pars <- c(new_pars,as.numeric(new_mat))
#   }
# 
#   new_pars <- c(new_pars,origin_pars[(ns*numCovIn + 1 + (ns - 1)*numCovTr*ns + numCovTr*ns):length(origin_pars)])
# 
#   newFull[[ns]] <- setpars(newFull[[ns]],new_pars)
#   newFull[[ns]] <- fit(newFull[[ns]], emcontrol= em.control(maxit = 1000,random.start = FALSE))
# }
# 
# save(newFull, file="data/newFull.RData")

```

```{r, include=F}

load("data/newFull.RData")

bestFull <- NULL
# Loop through each index and select the model with the higher log-likelihood
for(i in n_state_min : n_state_max) {
  if(logLik(fittedFullCtrst[[i]]) > logLik(newFull[[i]])) {
    bestFull[[i]] <- fittedFullCtrst[[i]]
  } else {
    bestFull[[i]] <- newFull[[i]]
  }
}

```

```{r, include=F}
BICs_simple <- BICs_Ctrl <- BICs_Coax <-  BICs_PrePost <-  BICs_Full <- list()

for(i in n_state_min:n_state_max) {
  BICs_simple[[i]] <- BIC(fittedSimple[[i]])
  BICs_Ctrl[[i]] <- BIC(bestCtrl[[i]])
  BICs_Coax[[i]] <- BIC(bestCoax[[i]])
  BICs_PrePost[[i]] <- BIC(bestPrePost[[i]])
  BICs_Full[[i]] <- BIC(bestFull[[i]])

}

# print("SIMPLE HMM models")
# BICs_simple
# 
# print("HMM with contrast on post Control vs others")
# BICs_Ctrl
# 
# print("HMM with contrast on post Coax vs others")
# BICs_Coax
# 
# print("HMM with Pre vs Post Contrast")
# BICs_PrePost
# 
# print("HMM with contrast levels Pre, Post Control and Post Coaxing")
# BICs_FullCtrst
# 
# print("HMM with contrast levels Pre, Post Control and Post Coaxing, new fitting")
# BICs_newFull

df_liks <- cbind(2:7,sapply(fittedSimple[2:7], logLik), sapply(bestCtrl[2:7], logLik), sapply(bestCoax[2:7], logLik), sapply(bestPrePost[2:7], logLik), sapply(bestFull[2:7], logLik))
df_liks

df_bics <- cbind(2:7,as.numeric(BICs_simple[2:7]), as.numeric(BICs_Ctrl[2:7]),as.numeric(BICs_Coax[2:7]), as.numeric(BICs_PrePost[2:7]), as.numeric(BICs_Full[2:7]))
df_bics

df_bics_full <- data.frame(nstate = 2:7,
                          logLik = unlist(lapply(bestFull[-1], logLik)),
                          unlist(lapply(bestFull[-1], function(x) attr(logLik(x), which="df"))),
                          AIC = unlist(lapply(bestFull[-1], AIC)),
                          BIC = unlist(lapply(bestFull[-1], BIC)))
```

```{r, include=F}

df_bics_full %>%
  `colnames<-` (c("Number of states", "logLikelihood", "df", "AIC", "BIC")) %>%
  mutate(across(all_of(c("logLikelihood", "df", "AIC", "BIC")), 
                ~format(., digits=4, nsmall=0, big.mark=","))) %>%
  knitr::kable(., escape = FALSE, booktabs = TRUE,
               caption = 'Table comparing HMM-full model fit for assumed number of latent states between 2 and 7') %>%
  kable_styling(latex_options="scale_down",font_size = 10)

```

```{r, include=FALSE}
library(knitr)
library(kableExtra)

# Function to make cell with lowest value bold 
bold_min <- function(x) {
  min_val = min(x)
  sapply(x, function(val) {
    if (val == min_val) {
      cell_spec(format(val, digits=4, nsmall=0, big.mark=","), "latex", bold = TRUE)
    } else {
      format(val, digits=4, nsmall=0, big.mark=",")
    }
  })
}

# Function to make cell with highest value bold 
bold_max <- function(x) {
  max_val = max(x)
  sapply(x, function(val) {
    if (val == max_val) {
      cell_spec(format(val, digits=4, nsmall=0, big.mark=","), "latex", bold = TRUE)
    } else {
      format(val, digits=4, nsmall=0, big.mark=",")
    }
  })
}

# Create a data frame for the models
models <- data.frame(
  Model = c("HMM-inv", "HMM-prepost", "HMM-coax", "HMM-ctrl", "HMM-full"),
  Pre_Intervention =  c(0, -1, -1, -1, -1),
  Post_Intervention = c(0, 1, 1, -1, 2),
  Pre_Control =       c(0, -1, -1, -1, -1),
  Post_Control =      c(0, 1, -1, 1, 1)
)

# Create a table
models %>%
  kable("pipe", caption = "Models and their Contrasts") %>%
  kable_styling()

```

```{r table-BICs-coax, include =F, warning = FALSE, echo=FALSE, ft.align="center", tab.playerId='table-BICs-coax', label='table-BICs-coax'}

df_liks <- as.data.frame(df_liks)

df_liks %>%
  `colnames<-` (c("Number of states", "HMM-inv", "HMM-ctrl", "HMM-coax", "HMM-prepost", "HMM-full")) %>%
  mutate_at(vars(-`Number of states`), bold_max) %>%
  knitr::kable(., escape = FALSE, booktabs = TRUE,
               caption = 'Table of likelihoods for each of the estimated HMM models for assumed number of latent states between 2 and 7') %>%
  kable_styling(latex_options="scale_down")

##### BICs ###########
# Ensure df_bics is a data frame
df_bics <- as.data.frame(df_bics)

df_bics %>%
  `colnames<-` (c("Number of states", "HMM-inv", "HMM-ctrl", "HMM-coax", "HMM-prepost", "HMM-full")) %>%
  mutate_at(vars(-`Number of states`), bold_min) %>%
  knitr::kable(., escape = FALSE, booktabs = TRUE,
               caption = 'Table of BICs for each of the estimated HMM models for assumed number of latent states between 2 and 7') %>%
  kable_styling(latex_options="scale_down")

```

We used hidden Markov models (HMMs) to further assess differences between the intervention and control condition in participants' reactions to the investor in the Repeated Trust Game. As in the models for the investor, these HMMs assume behavior is governed by latent states, with participants' switches between states now dependent on the investments made. We also allowed for differences between games and conditions in how investments govern state transitions: We fitted five main models which all regressed state transition probabilities onto investments, as well as on additional contrast-coded predictors for Condition and/or Game. In the most complex model (HMM-full), the transition probabilities were allowed to differ between all four combinations of Game and Condition. The HMM-coax model allowed differences between post-intervention and the other three conditions (pre-intervention, pre-control, post-control) treating these latter conditions as the same. Similarly, the HMM-ctrl model allowed differences between post-control and the other three conditions. The HMM-prepost model allowed differences between the first and second RTG. Finally, the HMM-inv model did not allow transition probabilities to differ between conditions or games, modelling them only as a function of investment. As the number of hidden states was unknown, we estimated models with 2 to 7 latent states for the most complex HMM-full model, and used the BIC to compare them. The best fitting HMM-full model according to the BIC had 6 latent states. Further details on the HMMs and estimation procedure are provided in the Supplementary Information.

```{r, include=FALSE}

llratio(bestFull[[6]], fittedSimple[[6]])
llratio(bestFull[[6]],bestCtrl[[6]])
llratio(bestFull[[6]],bestCoax[[6]])
llratio(bestFull[[6]],bestPrePost[[6]])

# llratio(bestFull[[6]],bestCtrl[[6]])
# llratio(bestFull[[6]],bestCoax[[6]])
# llratio(bestFull[[6]],bestPrePost[[6]])
```

Focusing on models with 6 latent states, likelihood ratio tests showed that the HMM-full model fits significantly better than HMM-ctrl ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCtrl[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$), HMM-coax ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestCoax[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$) and HMM-prepost ($\chi^2(60) = `r round(-2*as.numeric(logLik(bestPrePost[[6]])) - (-2*as.numeric(logLik(bestFull[[6]]))) , 2)`$, $p < .001$). As such, there is evidence that participants reacted differently to the investments in the four combinations of Condition and Game. The estimated state-dependent policy of trustee actions, according to the HMM-full model, is depicted in Figure \ref{fig:trusteeHMM}.A.

<!--    ORDER THE STATES    -->

```{r, include=FALSE}
tr_fdmod_6 <- order_mod_vtdgaus(bestFull[[6]])
```

```{r, include=FALSE}

state1 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[1]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

state2 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[2]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

state3 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[3]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

state4 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[4]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

state5 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[5]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

state6 <- dens(truncdiscgaus(seq(0.025,0.975,length=20),pstart=c(unlist(tr_fdmod_6@response[[6]][[1]]@parameters)), breaks = c(0, seq(0.05,0.95,length=19) , 1),  min=0, max=1))

returns <- seq(0.025,0.975,length=20)

response_best_mod <- as.data.frame(cbind(returns,state1,state2,state3,state4,state5, state6)) %>% 
  pivot_longer(cols=c("state1","state2","state3","state4","state5", "state6"),
                    names_to='Trustee_state',
                    values_to='probability')

```

```{r, include=F, fig.align="center"}

tr_policy <- list()
for(i in 1:nstates(tr_fdmod_6)) {
  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(tr_fdmod_6@response[[i]][[1]]@parameters))))  
}

perc_return <- seq(0,1, length=61)
df_tr_policy <- data.frame(trustee_state = rep(1:nstates(tr_fdmod_6), each=length(perc_return)),
                            perc_return = perc_return,
                            probability = unlist(tr_policy))

ggplot(df_tr_policy,                            
       aes(x = perc_return,
           y = probability,
           fill = factor(paste0("state ",trustee_state)))) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent trustee state') +
  xlab("Percentage return") +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

```{r, include=FALSE}

plot_state_transitions <- function(fmod, numCovIn, numCovTr) {

  #print(getpars(fmod))
  ns <- nstates(fmod)

  trans_prob <- data.frame(
    from = rep(paste0("from State ", 1:ns), each=100*ns),
    to = rep(paste0("State ", 1:ns), times=100),
    investment = rep(seq(0,20,length=100),each=ns),
    prob_pre = 0,
    prob_ctrl = 0,
    prob_coax =0
  )

  y0 <- matrix(0.0, ncol=ns, nrow=100)
  y1 <- matrix(0.0, ncol=ns, nrow=100)
  y2 <- matrix(0.0, ncol=ns, nrow=100)

  for(from in 1:ns) {
    pars <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)], ncol=numCovTr)
    #print(pars)

    for(to in 1:ns) {
        x <- trans_prob[trans_prob$from == paste0("from State ", from) & trans_prob$to == paste0("State ", to), "investment"]
        y0[,to] <- exp(pars[to,1] + pars[to,2]*x)
        y2[,to] <- exp(pars[to,1] + pars[to,2]*x + pars[to,3]*1 + pars[to,5]*x)
        y1[,to] <- exp(pars[to,1] + pars[to,2]*x + pars[to,4]*1 + pars[to,6]*x)
    }

    y0 <- y0/rowSums(y0)
    y1 <- y1/rowSums(y1)
    y2 <- y2/rowSums(y2)

    for(to in 1:ns) {
      trans_prob$prob_pre[trans_prob$from == paste0("from State ", from) & trans_prob$to == paste0("State ", to)] <- y0[,to]
      trans_prob$prob_ctrl[trans_prob$from == paste0("from State ", from) & trans_prob$to == paste0("State ", to)] <- y1[,to]
      trans_prob$prob_coax[trans_prob$from == paste0("from State ", from) & trans_prob$to == paste0("State ", to)] <- y2[,to]
    }
  }

  saveRDS(trans_prob, "data/trans_prob.RDS")

  return(trans_prob)
}

plot_state_transitions(tr_fdmod_6, numCovIn=2, numCovTr=6)

```

```{r,include=F}

trans_prob <- readRDS("data/trans_prob.RDS")

ctrl_pre_plot <- ggplot(trans_prob,aes(x=investment,y=prob_ctrl, colour = as.factor(to))) + 
  geom_line() + 
  geom_line(aes(y=prob_pre),linetype="dotted") +
  facet_wrap(~from) + 
  ylim(c(0,1)) + 
  ggtitle("Transition function pre- and post-control manipulation") + 
  labs(x = "Investment", y = "Transition probability", color='Transition to:') + 
  theme_bw()

coax_pre_plot <- ggplot(trans_prob,aes(x=investment)) + 
  geom_line(aes(y=prob_coax, colour = as.factor(to), linetype="Post-manipulation")) + 
  geom_line(aes(y=prob_pre, colour = as.factor(to), linetype="Pre-manipulation")) +
  facet_wrap(~from) + 
  ylim(c(0,1)) + 
  ggtitle("Transition function pre- and post-intervention manipulation") + 
  labs(x = "Investment", y = "Transition probability", color='Transition to:', linetype = "Line Type") + 
  scale_linetype_manual(values = c("Post-manipulation" = "solid", "Pre-manipulation" = "dotted")) +
  theme_bw()

```

```{r plotTransitionsDiscGaus, include=F, fig.cap="Transition function for the HMM-full trustee model. Each panel represents the state transitioned from. For instance the upper left panel represents all transition probabilities from state 1.  The colors of the lines within each panel indicate the state transitioned to. For instance, in the upper left panel, the red line represents the transition probability to state 1 as a function of the received investment. In this case this is also the probability of staying in state 1 since we are in the state 1 panel.  Solid lines show estimated transition probabillities post-manipulation while dotted lines show the same probabilities prior to the manipulation",fig.align="center", fig.width=6, fig.height = 7 }

grid.arrange(ctrl_pre_plot , coax_pre_plot , nrow=2, ncol=1)

```

```{r, include = FALSE}
set.seed(20221010)
# Get investor posterior states from model and add to data table

# mention we are using local decoding (refer to the book). 
predTrStates <- posterior(tr_fdmod_6, type="local")
avg_ret_df$TrState <-  factor(predTrStates, levels= c(1,2,3,4,5,6),labels=c("State 1","State 2","State 3","State 4","State 5", "State 6"))

library(dplyr)
temp_df <- avg_ret_df %>% dplyr::select(roundNum,investment,returns, next_investment, TrState)
```

```{r, include=FALSE}
postHMM <- ggplot(avg_ret_df %>% dplyr::filter(gameNum.f=="post")) +
  geom_bar(aes(x = roundNum, group = TrState, fill = TrState),  position = position_stack(reverse = TRUE)) + 
  facet_wrap(~factor(condition.f, levels = rev(levels(as.factor(condition.f))))) +  
  labs(x = "Round", fill='Posterior trustee state') +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r postStatesBestMod,  include=F, echo=FALSE, fig.cap = "Distribution of posterior trustee states post-manipulation by condition for all rounds, as estimated by the most likely posterior state in the best fitting HMM model (HMM-full) using a local decoding procedure. States are represented in increasing degrees of average percentage returns from the lowest (state 1) to the highest (state 6) return state. This figure characterises behavioral differences between conditions as the result of participants being in more pro-social (higher return) states in the intervention condition compared to the control condition both pre- and post-defection.", fig.align="center"}

postHMM

```

```{r, include= F}

props_df_5 <- avg_ret_df %>% 
  filter(gameNum.f=="post", roundNum== 5) %>% 
  group_by(condition.f,TrState) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

prop.test(c(8,20), c(159,159))

props_df_14 <- avg_ret_df %>% 
  filter(gameNum.f=="post", roundNum== 14) %>% 
  group_by(condition.f,TrState) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

prop.test(c(35,68), c(159,159))

props_df_13 <- avg_ret_df %>% 
  filter(gameNum.f=="post", roundNum== 13) %>% 
  group_by(condition.f,TrState) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

# compare proportions in state 1 in round 13 between conditions
prop.test(c(48,72), c(159,159))

# compare proportions in states 4 and 5  in round 13 between conditions
prop.test(c(48,18), c(159,159))

# props_df_12 <- avg_ret_df %>% 
#   filter(gameNum.f=="post", roundNum== 12) %>% 
#   group_by(condition.f,TrState) %>% 
#   summarise(n = n()) %>%
#   mutate(freq = n / sum(n))

```

Taking the best-fitting 6-state HMM-full model, we used a local decoding procedure to assign observations (participants' returns on trials) to latent states. The states are ordered by expected return, with state 1 having the lowest mean return and state 6 the highest. Figure \ref{fig:trusteeHMM}.B shows that participants were more likely to be in a lower return state in the control condition compared to the intervention condition, both pre and post defection. For instance, in round 5, state 1 was the most likely posterior state for only $5$% of participants in the intervention condition compared to $12$% in the control condition ($\chi^2(1) = 4.73, p = 0.03$). For the post-defection trial after the intervention (round 14), state 1 was the most likely state for only $15$% of participants in the intervention condition compared to $31$% in the control condition ($\chi^2(1) = 14.70, p < 0.001$). Whilst the posterior states indicate that the intervention was effective, a non-negligible proportion of participants in the intervention condition did not exhibit the coaxing behavior promoted by the intervention. Directly following the low investment in round 13, $17$% of participants in the intervention condition were assigned to state 1 with the lowest average returns highlighting individual differences in the effectiveness of the intervention.

# Discussion

Following a cognitive intervention, participants increased their returns without a corresponding change in emotional response, indicating the intervention's effectiveness in preserving trust despite a breach of cooperation. Those in the control condition returned similar proportions pre-defection and reduced their returns post-defection. An HMM analysis showed that those in the intervention condition were less prone to low-return states after defection. The findings suggest that cognitive interventions can promote resilience in cooperative behaviors following trust violations. This also validates the use of HMM-based artificial agents in economic games, effectively replicating investment and returns from human dyadic interactions. Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.

\newpage
```{r trusteeHMM, fig.cap="\\small{Panel A: the distribution of participants' percentage return for each of the latent states in the 6 state HMM-full model shows distinct policies centered around different return levels. The latent states are ordered by the mean of the discretized Gaussian representing the policy in that state, so higher numbered states can be considered more pro-social. B: This figure characterises behavioral differences between conditions as the result of participants being in more pro-social (higher return) states in the intervention condition compared to the control condition both pre- and post-defection. The distribution of posterior trustee states post-manipulation by condition for all rounds, as estimated by the most likely posterior state in the best fitting HMM model (HMM-full) using a local decoding procedure is visibly more weighted towards higher return states. As in Panel A, states here are represented in increasing degrees of average percentage returns from the lowest (state 1) to the highest (state 6) return state.}", echo=FALSE, message=FALSE, warning=FALSE, out.width='0.95\\linewidth', out.height='0.80\\textheight', fig.align='center'}
library(magick)

# Read in the images
image7 <- image_read("figures/trusteeHMMPlot.png")
image8 <- image_read("figures/postHMM.png")

image7 <- image_annotate(image7, text = "A", location = "+20+20", size = 70, color = "black", gravity = "northwest")
image8 <- image_annotate(image8, text = "B", location = "+20+20", size = 70, color = "black", gravity = "northwest")

# Combine image2 and image3 side-by-side
final_image <- c(image7, image8) %>% magick::image_append(stack = TRUE)

# Display the combined image
plot(final_image)
```
\newpage

That participants in the intervention condition generally send higher returns to the investor after the intervention is unlikely due to a  general learning effect unrelated to the intervention, as participants in the control condition did not increase their returns. Also, as participants in the intervention and control condition faced the same HMM investor, the higher post-intervention returns are not solely due to a difference in investor behavior. As the investor reacts to participants' returns, those who return more will generally see higher investments. But this is driven by the magnitude of their returns, not by a change in the strategy of the investor. Finally, as there were no differences between conditions in how participants rated the first and second HMM agent on attributes such as cooperativeness and trust, the increased returns are unlikely to be the result of a more favourable evaluation of the investor. Rather, we find it most likely that the increased returns are due to participants inferring from the intervention that pro-social and trustworthy behavior may generally motivate the investor to send high investments, which can provide more beneficial outcomes to them in the long run.

The effect of the intervention was not transferred to the Repeated Prisoner's Dilemma (RPD) game. There was no difference between the intervention and control condition in the rate of cooperation, whether before or after a preprogrammed defection by the artificial agent. As the prisoner's dilemma is a popular economic game, it is possible that participants had a strong prior commitment towards the strategy they would adopt, which was not overridden by the intervention. The RPD also involves much coarser actions (cooperate vs defect) than the finer-grained returns in the RTG. This makes it more difficult to observe more subtle effects of the intervention. As such, the RPD might not be the best choice of task to measure transfer. In any case, we can not rule out that the effects of the brief cognitive intervention, which explicitly focused on the RTG, are confined to the RTG. Future research will need to assess whether transfer to other games is possible.

Analysing participants' behavior with hidden Markov models, we found clear individual differences in how returns changed between the pre- and post-intervention RTG, which can be seen as a proxy for the effectiveness of the intervention. Some participants may not have been convinced that coaxing via high returns was a good way to establish cooperation and decided to reduce their returns in the second trust game. Their impulse to “punish” the other player for a defection may have been too strong to be overridden by the intervention. This was also evident from participants' replies to a question about whether they would change their behavior, just after receiving the intervention. Other participants responded to the intervention and increased their returns. Heterogeneity in response to treatment is common in psychiatry and related fields. Such heterogeneity may reflect the complex nature of mental health problems, which may be best viewed as complex systems involving interactions between neuro-computational processes and socio-environmental contexts evolving over time [@fried_moving_2017]. This view was used to justify computational psychiatry's difficulty in establishing differential and reliable predictors of likely treatment response [@hitchcock_computational_2022]. Here, we found heterogeneity in reaction to a relatively explicit intervention by a sample of participants from the general population. This suggests that the issue of variable treatment responses may result from the interaction of two sources of variability: the phenotyping of the disorder as well as the phenomenological aspects of the intervention itself. As such, a rigorous exploration of the determinants of inter-individual differences to an intervention in the general patient population is required.

Overall, we are encouraged that our brief cognitive intervention, consisting of reading a short text detailing a non-impulsive reaction to low investments, can lead to clearly differentiated behavior. In future studies, we aim to explore the effects of improved cognitive interventions to enhance cooperative behavior. We may enhance engagement by making the intervention more interactive and visually appealing, rather than the "dry" textual format used in the experiment. It would also be of interest to test such interventions with participants that suffer from an inability or unwillingness to repair relationships after an accidental breakdown of trust, such as people with Borderline Personality Disorder. The relative ease by which online interventions can be assigned, and the opportunity for people to test the effect of their behavior with artificial but human-like agents, may pave the way for efficient, low-cost, effective treatment programmes which may help a wide-variety of people overcome detrimental actions in social situations.

\pagebreak

# References